Comparing R packages for calculation of persistent homology

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised text
fixed gudhi capitalization

Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
Vietoris-Rips complices show rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
Interestingly, different factors determine memory usage in the two types of simplicial complex.
Whereas memory use in calculation of a Vietoris-Rips complex is dependent on point cloud size and feature dimension, memory use for an analogous calculation with an alpha complex is also dependent on the structure of the point cloud itself (Figure \ref{fig:mem2}).
Within the limited set of tested point cloud structures, the sphere requires the least memory.
Table \ref{tbl:limit} lists the maximum point cloud size for which persistent homology calculations were completed for each library.

@@

\begin{table}
\centering
\small
\begin{tabular}{|c|c|c|c|c|}
\hline
Feature & Ripser & Dionysus & Gudhi Rips & Gudhi Alpha \\
Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
\hline
1 & Not Reached & Not Reached & Not Reached & Not Reached \\ 
2 & Not Reached & 200 points & 280 points & Not Reached \\ 
3 & 300 points & 100 points & 125 points & Could Not Compute \\
4 & 100 points & 50 points & 75 points & Could Not Compute \\ 
\hline
\end{tabular}
\label{tbl:limit}
\caption{During the benchmarking, certain libraries were not able to calculate persistent homology for sufficiently complex point clouds. 
Limiting factors were point cloud number and feature dimension calculation. 
For Vietoris-Rips complices, Ripser had the best performing capacity among the three libraries. 
Alpha complexes were not computable for any point cloud greater than 3 dimensions, so dimensional features 3 and higher could not be computed.}
\end{table}

>>>>

Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
Vietoris-Rips complices show rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
Interestingly, different factors determine memory usage in the two types of simplicial complex.
Whereas memory use in calculation of a Vietoris-Rips complex is dependent on point cloud size and feature dimension, memory use for an analogous calculation with an alpha complex is also dependent on the structure of the point cloud itself (Figure \ref{fig:mem2}).
Within the limited set of tested point cloud structures, the sphere requires the least memory.
Table \ref{tbl:limit} lists the maximum point cloud size for which persistent homology calculations could be completed for each library.

@@

\begin{table}
\centering
\small
\begin{tabular}{|c|c|c|c|c|}
\hline
Feature & Ripser & Dionysus & GUDHI Rips & GUDHI Alpha \\
Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
\hline
1 & Not Reached & Not Reached & Not Reached & Not Reached \\ 
2 & Not Reached & 200 points & 280 points & Not Reached \\ 
3 & 300 points & 100 points & 125 points & Could Not Compute \\
4 & 100 points & 50 points & 75 points & Could Not Compute \\ 
\hline
\end{tabular}
\caption{During the benchmarking, certain libraries were not able to calculate persistent homology for sufficiently complex point clouds. 
Limiting factors were point cloud number and feature dimension calculation. 
For Vietoris-Rips complices, Ripser had the best performing capacity among the three libraries. 
Alpha complexes were not computable for any point cloud greater than 3 dimensions, so dimensional features 3 and higher could not be computed.}
\label{tbl:limit}
\end{table}
